Relationship between solar wind, $D_{st}$, and plasmasphere mass density on one-hour time scales

### 2015-06-30 paper.tex
Updated plots. Adding CC plot and updated F10.7 analysis

\twocolumn[
  \begin{@twocolumnfalse}
\maketitle
\hrule
\begin{abstract}
This paper looks at how various magnetosphere conditions behave around the onset of geomagnetic events. It confirms results from previous papers that a sudden drop in $D_{st}$ correlates with a spike in equatorial mass density, while adding a level of depth and specification to the applicability of their results. It compares data at 1-hour averages to data at 1-day averages, and briefly to data at 27-day averages to examine if the trend holds at varying time scales and finds that it does, so long as the data is pre-processed to linearly remove a large time scale $F_{10.7}$ dependence. 
By then looking at an hourly average of variables from 24 hours before event onset to 48 hours after, short timescale trends can be discerned.

@@

\cite{Takahashi2010} state that spikes in the Disturbance Storm Time ($D_{st}$) index coincide with with significant changes in $\rho_{eq}$ at an L-shell of 6.8$R_E$. For five storms over a 20 day period two had $\rho_{eq}$ spikes after the $D_{st}$ drop, two had $\rho_{eq}$ spikes before the drop, and one showed little change in $\rho_{eq}$. They then show an epoch analysis where $\rho_{eq}$ is seen to spike the day of a $D_{st}$ drop, using a daily average of 30 minute $\rho_{eq}$ and one hour $D_{st}$ measurements. Figure \ref{DailyAverages} show that the trends observed by \cite{Takahashi2010} hold true in this analysis as well.

@@

The parameters $\rho_{eq}$ and $F_{10.7}$ used in this work are from the data associated with \cite{Denton}, with data available from 1980-1991; all other variables are from \cite{Reconstruction} over the time range of 1972-2013, which are on a 1-hour time grid.  In cases where $\rho_{eq}$ was available from multiple Geostationary Operational Environmental Satellites (GOES) satellites at the same time, the value from GOES 7 was selected to avoid overlapping data points.  $\rho_{eq}$ is the inferred mass density based on the 3rd harmonic frequency of magnetic field measurements.  The smallest cadence for $\rho_{eq}$ values is 10 minutes.  To compute an hourly average over the same time range as the solar wind parameters, the median of all values in a given hourly range was used.  Fill values were used for hours when no measurements were available.

@@

\subsection{D$_{st}$ Events}
Two event indicators are looked at in this study. The first is looking for a drop in DST below the threshhold of $-40nT$ specified in \cite{Takahashi2010}, dubbed the "onset", and then considering the timeframe an event until $D_{st}$ passes back above the $-40nT$ threshold. This method finds 669 such periods between May 1983 and August 1991 with an average duration of 9 hours and a median duration of 3 hours. Figure \ref{Storms} shows the average values of all events over a window of 24 hours before onset and 48 hours after. Figure \ref{Storms}a shows events selected by looking for event onsets where $D_{st}$ crossed a threshold of $-40nT$. The dashed red lines indicate plus and minus one standard deviation of values from all events that went into the average. The final plot in the stack shows both $\rho_{eq}$ and a bar plot of how many valid data points went into the $\rho_{eq}$ average. Since $\rho_{eq}$ comes from a sparser dataset, it has less valid points contributing to the averages than the other variables in the stack. A subset of longer duration events will be looked at later.

@@

Unforunately there are no events in this time frame that are longer than 12 hours with $D_{st}$ minima lower than $-80nT$ that have existing mass density data around onset, so an analysis of this particular relationship can't be made.

@@

@@

\begin{figure}[htp]
\centering
\includegraphics[scale=0.7]{paperfigures/alldata.eps}
\caption{Overview of data used in this study. The top four panels are from \cite{Reconstruction} and the bottom panel is based on \cite{Denton} after interpolation and averaging described in the text.}
\label{AllData}
\end{figure}
\clearpage

@@

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.5]{paperfigures/stormavs-dd12.eps}
\includegraphics[scale=0.5]{paperfigures/stormavs-d80.eps}
\caption{(a) Same as \ref{fig:MassStorm}(a) except with constraint that $D_{st}$ stayed below -40~nT for at least 12~hours. (b) Same as \ref{fig:MassStorm}(b) except for constrain that $D_{st}$ crossed below $-80 nT$.}
\label{Mspec}
\end{figure}
\clearpage

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.5]{paperfigures/stormavs-md12.eps}
\includegraphics[scale=0.5]{paperfigures/stormavs-m70.eps}
\caption{(a) Same as \ref{fig:MassStorm}(a) except with constraint that $\rho_{eq}$ stayed above 50~nT for at least 12~hours. (b) Same as \ref{fig:MassStorm}(b) except for constrain that mass density crossed above $70 amu/cm^3$.}
\label{Dspec}
\end{figure}
\clearpage

>>>>

\twocolumn[
  \begin{@twocolumnfalse}
\maketitle
\hrule
\begin{abstract}
This paper looks at how various magnetosphere conditions behave around the onset of geomagnetic events. It confirms results from previous papers that a sudden drop in $D_{st}$ correlates with a spike in equatorial mass density, while adding a level of depth and specification to the applicability of their results. It compares data at 1-hour averages to data at 1-day averages, and briefly to data at 27-day averages to examine if the trend holds at varying time scales and finds that it does, so long as the data is pre-processed to linearly remove a large time scale $F_{10.7}$ dependence. 
By then looking at an hourly average of parameters from 24 hours before event onset to 48 hours after, short timescale trends can be discerned.

@@

\cite{Takahashi2010} state that spikes in the Disturbance Storm Time ($D_{st}$) index coincide with with significant changes in $\rho_{eq}$ at an L-shell of 6.8$R_E$. For five storms over a 20 day period two had $\rho_{eq}$ spikes after the $D_{st}$ drop, two had $\rho_{eq}$ spikes before the drop, and one showed little change in $\rho_{eq}$. They then show an epoch analysis where $\rho_{eq}$ is seen to spike the day of a $D_{st}$ drop, using a daily average of 30 minute $\rho_{eq}$ and one hour $D_{st}$ measurements.

@@

The parameters $\rho_{eq}$ and $F_{10.7}$ used in this work are from the data associated with \cite{Denton}, with data available from 1980-1991; all other parameters are from \cite{Reconstruction} over the time range of 1972-2013, which are on a 1-hour time grid.  In cases where $\rho_{eq}$ was available from multiple Geostationary Operational Environmental Satellites (GOES) satellites at the same time, the value from GOES 7 was selected to avoid overlapping data points.  $\rho_{eq}$ is the inferred mass density based on the 3rd harmonic frequency of magnetic field measurements.  The smallest cadence for $\rho_{eq}$ values is 10 minutes.  To compute an hourly average over the same time range as the solar wind parameters, the median of all values in a given hourly range was used.  Fill values were used for hours when no measurements were available.

@@

\subsection{Previous Results}
Figure \ref{DailyAverages} show that the trends observed by \cite{Takahashi2010} for one day averages only seems to hold true for the one-day averaged data, and only after the $F_{10.7}$ dependence is removed. The correlation seen between $log(\rho_{eq\_27d})$ and $F_{10.7\_27d}$ isn't quite as good in this data as in \cite{Takahashi2010}, but shows the same trend as seen in Figure \ref{ccplot}.

\subsection{D$_{st}$ Events}
Two event indicators are looked at in this study. The first is looking for a drop in DST below the threshhold of $-40nT$ specified in \cite{Takahashi2010}, dubbed the "onset", and then considering the timeframe an event until $D_{st}$ passes back above the $-40nT$ threshold. This method finds 669 such periods between May 1983 and August 1991 with an average duration of 9 hours and a median duration of 3 hours. Figure \ref{Storms} shows the average values of all events over a window of 24 hours before onset and 48 hours after. Figure \ref{Storms}a shows events selected by looking for event onsets where $D_{st}$ crossed a threshold of $-40nT$. The dashed red lines indicate plus and minus one standard deviation of values from all events that went into the average. The final plot in the stack shows both $\rho_{eq}$ and a bar plot of how many valid data points went into the $\rho_{eq}$ average. Since $\rho_{eq}$ comes from a sparser dataset, it has less valid points contributing to the averages than the other parameters in the stack. A subset of longer duration events will be looked at later.

@@

Unfortunately there are no events in this time frame that are longer than 12 hours with $D_{st}$ minima lower than $-80nT$ that have existing mass density data around onset, so an analysis of this particular relationship can't be made.

@@

\begin{figure}[htp]
\centering
\includegraphics[scale=0.7]{paperfigures/alldata.eps}
\caption{Overview of data used in this study. The top four panels are from \cite{Reconstruction} and the bottom panel is based on \cite{Denton} after interpolation and averaging described in the text. Dashed lines indicate default event cutoff thresholds.}
\label{AllData}
\end{figure}
\clearpage

@@

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.5]{paperfigures/stormavs-dst-day-nof107.eps}
\includegraphics[scale=0.5]{paperfigures/stormavs-dst-27day-nof107.eps}
\caption{One-day and 27-day averaged $D_{st}$ storms with $F_{10.7}$ dependence removed}
\label{DailyAveragesNoF107}
\end{figure}
\clearpage

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.7]{paperfigures/ccplot.eps}
\caption{Correlation between $log(\rho_{eq\_27d})$ and $F_{10.7\_27d}$ }
\label{ccplot}
\end{figure}
\clearpage

@@

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.5]{paperfigures/stormavs-dd12.eps}
\includegraphics[scale=0.5]{paperfigures/stormavs-d80.eps}
\caption{(a) Same as \ref{fig:Storms}(a) except with constraint that $D_{st}$ stayed below -40~nT for at least 12~hours. (b) Same as \ref{fig:Storms}(b) except for constrain that $D_{st}$ crossed below $-80 nT$.}
\label{Mspec}
\end{figure}
\clearpage

\begin{figure}[htp!]
\centering
\includegraphics[scale=0.5]{paperfigures/stormavs-md12.eps}
\includegraphics[scale=0.5]{paperfigures/stormavs-m70.eps}
\caption{(a) Same as \ref{fig:MassStorm}(a) except with constraint that $\rho_{eq}$ stayed above 50~nT for at least 12~hours. (b) Same as \ref{fig:Storms}(b) except for constrain that mass density crossed above $70 amu/cm^3$.}
\label{Dspec}
\end{figure}
\clearpage
